---
title: "Homework 3"
author: "Caroline Ledbetter"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 5
    fig_width: 4
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = '../')
```


```{r setup, include=FALSE}
library(foreign)
library(gbm)
library(reshape2)
library(pROC)
TBI <- read.spss('../DATA/TBI.sav')
TBI <- as.data.frame(TBI)
```
git repository: https://github.com/ledbettc/BIOS6645

```{r Q1a, include=F}
TBI$d.unfav <- as.logical(TBI$d.unfav)
TBI$d.motor <- factor(TBI$d.motor)
TBI_train <- subset(TBI, trial == "Tirilazad US", 
                  select = c(d.unfav, d.motor, d.pupil, age))
TBI_test <- subset(TBI, trial == "Tirilazad International", 
                  select = c(d.unfav, d.motor, d.pupil, age))
set.seed(1234) 
gbm1 <- gbm(d.unfav ~ d.motor + d.pupil + age, data = TBI_train,  
            distribution = "bernoulli",
            n.trees = 10000, interaction.depth = 2, cv = 10)
summary(gbm1)
print(gbm1)
best.iter <- gbm.perf(gbm1, method = "cv")
pred <- list()
pred$gbmtrain <- predict(gbm1, data = TBI_train, type = 'response')

roc_gbm <- roc(TBI_train$d.unfav, pred$gbmtrain)
plot(roc_gbm, print.thres = 'best', print.thres.best.method = "youden", 
     print.auc = T, main = 'ROC for GBM')
youd_gbm <- coords(roc_gbm, x = "best", best.method = "youden", 
                   ret = "threshold")
pred$gbmtrain2 <- ifelse(pred$gbmtrain <= youd_gbm, F, T)
acc <- list()
acc$gbmtrain <- sum(pred$gbmtrain2 == TBI_train$d.unfav)/nrow(TBI_train)
pred$gbmtest <- predict(gbm1, data = TBI_test, type = 'response')
pred$gbmtest2 <- ifelse(pred$gbmtest <= youd_gbm, F, T)
acc$gbmtest <- sum(pred$gbmtest2 == TBI_test$d.unfav)/nrow(TBI_test)
```

1.  
  a.   
    ii. N = `r best.iter` trees  


```{r plotsq1}
#Partial dependence plots - assess functional form suggested by results of gbm
par(mfrow =c(2,2))
summary.gbm(gbm1, n.trees=best.iter, 
            method=relative.influence) # variable importance plot
# visualize what gbm uncovers as functional form
plot(gbm1, i="d.motor", n.trees=best.iter) 
plot(gbm1, i="d.pupil", n.trees=best.iter)
plot(gbm1, i="age", n.trees=best.iter)
par(mfrow =c(1,1))
```

1.   
  a.   
    iv. I chose a thresholf of `r round(youd_gbm, 2)` based on the RUC curve using youden's     method with `r best.iter` trees.   
    vi. For the training data set (the US data) the predictive accuracy is 
        `r round(acc$gbmtrain, 2)`% and for the validation data set 
        (the international data)
        it is `r round(acc$gbmtest, 2)`.%   
      
```{r 1b, include = F}
glm1 <- glm(d.unfav ~ d.motor + d.pupil + age, data = TBI_train, 
            family = 'binomial')

pred$glmtrain <- predict(glm1, data = TBI_train, type = 'response')
pred$glmtrain2 <- ifelse(pred$glmtrain <= youd_gbm, F, T)
acc$glmtrain <- sum(pred$glmtrain2 == TBI_train$d.unfav)/nrow(TBI_train)

```
  
  b.  
    i. d.motor>d.pupil>age  
    ii. For the training data set (the US data) the predictive accuracy is 
        `r round(acc$glmtrain, 2)`% and for the validation data set 
        (the international data)
        it is round(acc$glmtest, 2).%
    iii. Yes. 
            
```{r show-code, ref.label = knitr::all_labels(), echo = TRUE, eval=FALSE}

```