---
title: "Homework 3"
author: "Caroline Ledbetter"
output:
  pdf_document:
    fig_caption: yes
    fig_height: 5
    fig_width: 4
---
```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
knitr::opts_knit$set(root.dir = '../')
```


```{r setup, include=FALSE}
library(foreign)
library(gbm)
library(reshape2)
library(pROC)
library(car)
library(mice)
library(VIM)
TBI <- read.spss('../DATA/TBI.sav')
TBI <- as.data.frame(TBI)
```
git repository: https://github.com/ledbettc/BIOS6645

```{r Q1a, include=F}
TBI$d.unfav <- as.logical(TBI$d.unfav)
TBI$d.motor <- factor(TBI$d.motor)
TBI_train <- subset(TBI, trial == "Tirilazad US", 
                  select = c(d.unfav, d.motor, d.pupil, age))
TBI_test <- subset(TBI, trial == "Tirilazad International", 
                  select = c(d.unfav, d.motor, d.pupil, age))
set.seed(1234) 
gbm1 <- gbm(d.unfav ~ d.motor + d.pupil + age, data = TBI_train,  
            distribution = "bernoulli",
            n.trees = 10000, interaction.depth = 2, cv = 10)
summary(gbm1)
print(gbm1)
best.iter <- gbm.perf(gbm1, method = "cv")
pred <- list()
pred$gbmtrain <- predict(gbm1, data = TBI_train, type = 'response')
pred$gbmtest <- predict(gbm1, newdata = TBI_test, type = 'response')

roc_gbm <- roc(TBI_train$d.unfav, pred$gbmtrain)
plot(roc_gbm, print.thres = 'best', print.thres.best.method = "youden", 
     print.auc = T, main = 'ROC for GBM')
AUC <- list()
AUC$gbmtrain <- auc(roc_gbm)
AUC$gbmtest <- auc(roc(TBI_test$d.unfav, pred$gbmtest))

```

1.  
  a.   
    ii. N = `r best.iter` trees  


```{r plotsq1}
#Partial dependence plots - assess functional form suggested by results of gbm
par(mfrow =c(2,2))
summary.gbm(gbm1, n.trees=best.iter, 
            method=relative.influence) # variable importance plot
# visualize what gbm uncovers as functional form
plot(gbm1, i="d.motor", n.trees=best.iter) 
plot(gbm1, i="d.pupil", n.trees=best.iter)
plot(gbm1, i="age", n.trees=best.iter)
par(mfrow =c(1,1))
```

1.   
  a.
    iii.  Admission motor skill is the most predictive. Reactive pupils and age appear to have a somewhat linear form. Admission motor skills is perhaps non-linear. 
    iv. I chose a model with all 3 predictors and `r best.iter` trees.   
    v.  The admission motor score is most important followed by the pupil 
        reactivity then age. Age is much less important than the other two.  
    vi. For the training data set (the US data) the AUC is 
        `r round(AUC$gbmtrain, 2)`% and for the validation data set 
        (the international data)
        it is `r round(AUC$gbmtest, 2)`.%   
      
```{r 1b}
glm1 <- glm(d.unfav ~ d.motor + d.pupil + age, data = TBI_train, 
            family = 'binomial')

pred$glmtrain <- predict(glm1, newdata = TBI_train, type = 'response')
pred$glmtest <- predict(glm1, newdata = TBI_test, type = 'response')

roc_glm <- roc(TBI_train$d.unfav, pred$glmtrain)
AUC$glmtrain <- auc(roc_glm)
AUC$glmtest <- auc(roc(TBI_test$d.unfav, pred$glmtest))

Anova(glm1, type="III", test = "LR")
```
  
  b.  
    i.  Using liklihood ratio tests: d.motor>age>d.pupil  
    ii. For the training data set (the US data) the predictive accuracy is 
        `r round(AUC$glmtrain, 2)`% and for the validation data set 
        (the international data)
        it is `r round(AUC$glmtest, 2)`.%   
    iii.Yes, the accuracy is better on the training data set than the test set 
        indicating some overfitting.  
        
```{r 2a}
TBI_miss <- read.csv(file = '../DATA/tbi_miss_2.csv')
TBI_miss_train <- subset(TBI_miss, trial == "Tirilazad US", 
                         select = c(d_unfav, d_pupil, d_motor, age))
md.pattern(TBI_miss_train)
aggr(TBI_miss_train, numbers=TRUE, sortVars=TRUE)
```

```{r 2acont, include = F}
TBI_mI <- mice(TBI_miss_train, m=5, maxit=50, meth='rf', 
               seed=444)
```

2. a. There are 27 individuals missing age, because only age is missing there are MAR missing on x(age). I don't have reason to suspect they are MNAR but can't be sure. 

```{r 2b}
glm_mI <- lapply(1:5, function(x){
  df <- TBI_miss_train
  df$age[row.names(df) %in% 
           row.names(TBI_mI$imp$age)] <- TBI_mI$imp$age[, x]
  glm <- glm(d_unfav ~ d_motor + d_pupil + age, data = df, 
            family = 'binomial')})
lapply(glm_mI, summary)
preds <- lapply(glm_mI, predict, data = TBI_miss_train, 
                type = 'response')
AUCs <- lapply(preds, function(x){
  auc(roc(TBI_miss_train$d_unfav, x))})
AUCs
```


```{r show-code, ref.label = knitr::all_labels(), echo = TRUE, eval=FALSE}

```